---
title: "PS1"
author: "ChenYen Liu_Yuzhu Liu_Ziyue Wang"
date: "2023-01-27"
output: md_document
---


```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rsample)  # for creating train/test splits
library(caret)
library(modelr)
library(parallel)
library(foreach)
library(patchwork)
library(tidyverse)
library(mosaic)
library(ggplot2)
library(data.table)
library(rsample)
library(caret)
library(modelr)
library(parallel)
library(foreach)
library(latexpdf)
```
# Question1 - Flights at ABIA

```{r, echo=FALSE,message=FALSE, warning=FALSE}
#read data
#ABIA<-read_csv('/Users/hiziyue/Desktop/2023 Spring/RA/ABIA.csv')

ABIA<-read_csv('/Users/yuzhuliu/Desktop/Data Mining/PS1/ABIA.csv')

by_monthly_destination = ABIA %>% 
  group_by(Month, Dest) %>% 
  summarize(count = n(),
            mean_DepDelay = mean(DepDelay, na.rm=TRUE))

ave_1 <- aggregate(mean_DepDelay ~ Month, by_monthly_destination, mean)
```

### Question a: What's the best time of year to fly to minimize delays?
```{r, echo=FALSE,message=FALSE, warning=FALSE}
ggplot(ave_1) + geom_col(aes(x=factor(Month), y=mean_DepDelay))
```
### Question b: Does the best time to fly to min delays change by destination?
```{r, echo=FALSE,message=FALSE, warning=FALSE}
#choose top 10 popular destinations
install.packages("plyr",repos = "http://cran.us.r-project.org")
library(plyr)
count(ABIA, "Dest")
c2 <- count(ABIA, "Dest")
by_monthly_destination1 <- merge(x=by_monthly_destination, y=c2, by="Dest", all.x = TRUE)

library(dplyr)
by_monthly_destination1 <- filter(by_monthly_destination1, freq > 1350)
by_monthly_destination1 <- filter(by_monthly_destination1, freq <10000)
```
```{r, echo=FALSE,message=FALSE, warning=FALSE}
#ggplot
library(ggplot2)
ggplot(by_monthly_destination1) + 
  geom_col(aes(x=factor(Month), y=mean_DepDelay)) + 
  facet_wrap(~Dest)
```
# 2) Wrangling the Olympics

## (2.A) What is the 95th percentile of heights for female competitors across all Athletics events

95th percentile of heights for female competitors across all Athletics events is 186.
```{r echo=FALSE}

Olym<-read.csv("/Users/yuzhuliu/Desktop/Data Mining/PS1/olympics_top20.csv")

OlymF<-Olym%>%
 subset(sex %in% c("F"))

  
  result <- quantile(OlymF$height, c(0.95))
#print(result)

```
## (2.B) Which single women's event had the greatest variability in competitor's heights across the entire history of the Olympics, as measured by the standard deviation?

Rowing Women's Coxed Fours has the greatest variability, with standard derivation equals to 10.9
 
```{r echo=FALSE}

Olymevent<-OlymF%>%
  group_by(event) %>%
  summarize(avg_height = mean(height),
            sd_height  = sd(height))%>%
 arrange(desc(sd_height))

#Olymevent %>% arrange(desc(sd_height))
  
```
## (2.C) How has the average age of Olympic swimmers changed over time? Does the trend look different for male swimmers relative to female swimmers? 

The trend between female and male swimmers is different before around 1935,where age of male swimmers decreases and then increases, and then decreases, and age of female swimmers slightly increases.

After 1950, age of all swimmers has an increasing trend.


```{r echo=FALSE,fig.height = 10}

OlymSwim<-Olym%>%
 subset(sport %in% c("Swimming"))%>%
  group_by(year) %>%
  summarize(avg_height = mean(height))

p3<-ggplot(OlymSwim) + 
  geom_line(aes(x=year, y=avg_height))+
  labs(title="Total swimmer average age change")

OlymSwimF<-Olym%>%
   subset(sex %in% c("F"))%>%
  subset(sport %in% c("Swimming"))%>%
  group_by(year) %>%
  summarize(avg_height = mean(height))

OlymSwimM<-Olym%>%
   subset(sex %in% c("M"))%>%
  subset(sport %in% c("Swimming"))%>%
  group_by(year) %>%
  summarize(avg_height = mean(height))


p4<-ggplot() + 
  #geom_line(data = OlymSwim, aes(x=year, y=avg_height, color="Total"))+
  geom_line(data = OlymSwimF, aes(x=year, y=avg_height,color="Female")) +
  geom_line(data = OlymSwimM, aes(x=year, y=avg_height, color="Male")) +
  labs(title="Swimmer average age change")+
  theme(legend.position="bottom", legend.direction="horizontal")

p3/p4
   


```
## Question 3


```{r echo=FALSE}
sclass <- read.csv("/Users/yuzhuliu/Desktop/Data Mining/PS1/sclass.csv")
```
```{r echo=FALSE, include=FALSE}
trim_350 <- filter(sclass, trim == "350")
trim_65AMG <- filter(sclass, trim == "65 AMG")
trim_350_split <- initial_split(trim_350, prop = 0.8)
trim_65AMG_split <- initial_split(trim_65AMG, prop = 0.8)
trim_350_training <- training(trim_350_split)
trim_350_testing <- testing(trim_350_split)
trim_65AMG_training <- training(trim_65AMG_split)
trim_65AMG_testing <- testing(trim_65AMG_split)
```
```{r echo=FALSE, include=FALSE}
k_grid = c(2, 4, 6, 8, 10, 15, 20, 25, 30, 35, 40, 45,
           50, 60, 70, 80, 90, 100, 125, 150, 175, 200)
grid_350 = foreach(k = k_grid, .combine='rbind') %do% {
  models_350 = knnreg(price ~ mileage,
                  data=trim_350_training, k=k)
  errs_350 = modelr::rmse(models_350, data=trim_350_testing)
  c(k=k, err_350 = mean(errs_350))
} %>% as.data.frame


ggplot(grid_350) + 
  geom_point(aes(x=k, y=err_350)) + 
  scale_x_log10() +
  ggtitle("RMSE under different K for trim=350")
```
From the above plot, when k = 10, we can get the minimal RMSE.

```{r echo=FALSE, include=FALSE}
model_350 = knnreg(price ~ mileage, data=trim_350_training, k=10)
trim_350_testing = trim_350_testing %>%
  mutate(price_pred = predict(model_350, trim_350_testing))
origin_350 <- ggplot(data = trim_350_testing) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.2)
origin_350 + geom_line(aes(x = mileage, y = price_pred), color='darkred', size=1.5) + ggtitle("Predict value for trim=350")
```
```{r echo=FALSE, include=FALSE}
k_grid = c(2, 4, 6, 8, 10, 15, 20, 25, 30, 35, 40, 45,
           50, 60, 70, 80, 90, 100, 125, 150, 175, 200)
grid_AMG = foreach(k = k_grid, .combine='rbind') %do% {
  models_AMG = knnreg(price ~ mileage,
                  data=trim_65AMG_training, k=k)
  errs_AMG = modelr::rmse(models_AMG, data=trim_65AMG_testing)
  c(k=k, err_AMG = mean(errs_AMG))
} %>% as.data.frame


ggplot(grid_AMG) + 
  geom_point(aes(x=k, y=err_AMG)) + 
  scale_x_log10() +
  ggtitle("RMSE under different K for trim=65 AMG")
```
From the above plot, when k = 35, we can get the minimal RMSE.
```{r echo=FALSE, include=FALSE}
model_AMG = knnreg(price ~ mileage, data=trim_65AMG_training, k=10)
trim_65AMG_testing = trim_65AMG_testing %>%
  mutate(price_pred = predict(model_AMG, trim_65AMG_testing))
origin_AMG <- ggplot(data = trim_65AMG_testing) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.2)
origin_AMG + geom_line(aes(x = mileage, y = price_pred), color='darkblue', size=1.5) + ggtitle("Predict value for trim = 65 AMG")
```
From the above two plots, under the same mileage, trim = 65AMG has a higher value.
This may because people have a preference for trim 65AMG or people have a higher demand for this sub-model.